A robust framework for variable extrusion width concentric toolpath generation

### 2019-07-26 4_method.tex
reformulate: significant => marked
because we filter unmarked regions

\begin{enumerate}
\item Compute a skeletonization of the outline shape based on the Voronoi diagram.
\item Use a significance measure to mark portions of the skeleton which lie in the center.
\item Determine the bead count values in the marked regions and handle transitions to different bead count values.
\item Generate junctions along non-marked bones of the skeleton and connect the junctions together to form the toolpaths.
\end{enumerate}

@@

\subsubsection{Significance measure}\label{sec:significance_measure}
Places where the naive concentric toolpathing strategy would create large underfilling and overfilling areas are locations which intuitively look like `the center' of the shape.
This can be formalized by looking at the significance measure knows as the \emph{bisector angle}.
The bisector angle $\alpha$ is the angle $\angle{p_ovp_1}$ between a location $v$ on a bone of the skeleton and its two supporting polygon points $p_0$ and $p_1$. \cite{attali1996modeling}
For a polygon with a pointy wedge area of an angle $\beta$, we have $\alpha = 180\degree - \beta$, which corresponds to overfill areas and underfill areas the size of $\nicefrac12 w^2 \left( \nicefrac14 \tan ( \alpha / 2) - \alpha / 2 \right)$ when filled using a naive strategy of constant bead width $w$.
See \cref{naive_overfill_underfill}.
The bisector angle is therefore an exact measure of the amount of overfill and underfill.
Contrary to related literature we will \emph{keep} the insignificant regions of the skeleton.
Our framework will employ a different toolpathing strategy to prevent overfill and underfill areas if they would be so large that $\alpha > \alpha_\text{max}$.

@@

A computationally efficient way to compute $\alpha$ can be obtained by looking at the ratio between feature radius $R$ and the Euclidean distance.
if $ | R(v_1) - R(v_0) | / |v_1 - v_0| >  \cos(\alpha_\text{max} / 2)$ then $\alpha > \alpha_\text{max}$.
See \cref{distance_based_angles}.
This ratio can readily be applied to simple skeleton segments to determine whether they are significant.
For skeletal segments generated by a polygon vertex at $(0,d)$ and a line segment through $(0,0)$ and $(0,\epsilon)$ or another vertex $(0,0)$, we can determine the significant portion by evaluating $\frac{\partial R}{\partial x} > \cos(\alpha_\text{max} / 2)$.
For parabolas we introduce extra points in the discretization at the two bounds where $| x | = d  / \tan(\alpha_\text{max} / 2)$
and for bones with vertex-vertex support we introduce discretization points at the two bounds where $| x | = \nicefrac12 d  / \tan(\alpha_\text{max} / 2)$.
This ensures that the discretization doesn't alter the extent and existance of significant regions.

@@

\subsubsection{Bead counts}
In order to apply the beading strategy to the VQ
we assign each significant node $v$ of the skeleton a bead count $v.B^*$ which is initialized to $b(R(v))$, but it is modified in order to prevent discontinuities and remaining gaps.

We also assign a bead count $v.B^*=b(R(v))$ to all nodes $v$ which are local maxima, i.e. the $v$ for which $\forall v_\text{n} \in v.\text{neighborhood} : R(v) > R(v_\text{n})$. 
This prevents the single overfill or underfill region in non-significant regions.
See \cref{rounded_vs_unrounded}.

\subsubsection{Transitioning}
In order to prevent discontinuities in the toolpaths around locations where the bead count changes, we need a way to smoothly transition to a different number of beads around locations where the 3D model has a changing radius.
See \cref{transitions}.
The transition will be placed anchored on each location $v$ where $R(v) = \nicefrac12 b^{-1}(n+1)$ for any lower bead count $n$.
At the start $v_0$ and end $v_1$ of the transition we introduce a new node in the skeleton and connect it to the outline as per the VQ requirement.
We then assign $v_0.B^* = n$ and $v_1.B^* = n + 1$.
If there is any node $v_x$ in between $v_0$ and $v_1$ we assign it a fractional bead count by linear interpolation: $v_x.B^* = n + D(v_0, v_x)/t(n)$, where $D(.)$ is the Euclidean distance measured along the significant bone(s).
See \cref{distance_rounding_transition}.

@@

\subsubsection{Transitioning to insignificance}
In significant regions of the VQ we employ the approach of applying the optimal bead counts and employing transitioning, while in insignificant regions the beading is only governed by the beading of locations with locally optimal feature radius.
Significant bones always form chains and don't branch out because of the angle constraint on $\alpha_\text{max}$.
At the end of such chain $v$ we may need to transition from the former approach to the latter.
This only happens when the feature radius continues increasing: if $\exists v_n \in v.\text{neighborhood}: R(v) < R(v_n) \land | R(v) - R(v_n) | / |v - v_n| >  \cos(\alpha_\text{max} / 2)$ then the bead count assigned by the significant region doesn't fit with a beading which would follow from a local optimum at $v_n$ or some farther node.
See \cref{transition_to_insignificance_problem}.

@@

\begin{figure}
\centering
\includegraphics[width=.5\columnwidth]{sources/method/transition_to_insignificance_problem.pdf}
\caption{
Problems at significance boundary regions highlighted with red ovals.
The beading approach from the significant region (blue) results in toolpaths (black) which don't line up with the approach following from the beading which results from the bead count at the center bottom.
}
\label{transition_to_insignificance_problem}
\end{figure}

The transition length of these transitions to insignificance is determined by the fractional difference in bead count between the two beadings:
if $v.B^* < n$ then $t = t(n) . (R(v) - v.L_n)$
otherwise $t = t(n) . (1 - (R(v) - v.L_n))$,
where $n = \argmax_i v.L_i < R(v)$.
At a distance $t$ from $v$ along significant bones we introduce an new node $v_0$ for the end of the transition and again we connect it to the outline as per the VQ requirements.

We modify $v.B^*$ such that it coincides with the beading propagated to $v$ from some farther local optimum or another significant region.
(This propagation is explained in \cref{section_toolpath_extraction}.)
We assign $v.B^* = n + 1 - \epsilon$, so that the transition switches from the significant beading to the beading of the insignificant approach entirely.
For nodes $v_x$ in between $v_0$ and $v$ we interpolate $v_x.B^*$ again linearly based on the Euclidean distance along the bones in between $v$ and $v_0$.

@@

The angle requirement ensures that there is never a 3-way junction of significant bones.

@@

Now that we have the skeleton set up, we determine the locations along the bones where the vertices of the toolpath polygons should be: the \emph{junctions}.
For significant regions we have already inroduced joints around the transitions which serve as those locations.
For insignificant regions we can simply apply the beading strategy in order to find the toolpath verts.

This needs to be stable against small perturbations in the outline polygon.
See \cref{heterogeneous_joint_generation}.
For any sequence on insignificant bones from the outline polygon inward we define the stable end point as the first vertex which is connected to a significant VQ segment.
Based on the feature radius $R$ there we place junctions along the insignificant bones following the beading strategy.

@@

\subsubsection{Toolpath extraction algorithm}
After the beading strategy has been applied we have a beading associated with each significant node as well as each node with locally optimal $R$.
In order to determine the junction locations of quads which don't contain any such node, we propagate the beading information down.
We therefore sort all quads on their highest $R$-value and process the high $R$ quads first.
The beading information is then passed down along the bones of the quad to lower $R$ nodes if those nodes don't already have a beading associated.

@@

When all extrusion segments are generated, we combine them into semi-continuous polyline toolpaths.
These polylines generally form closed polygons, but at significant regions of the VQ the toolpaths consist of single bead polyline toolpaths.
See \cref{single_bead_strategy}.
These single bead polylines can connect to polygonal toolpaths and they can also meet each other at 3-way junctions.
We determine the printing order for all segments of a given bead index by minimizing the travel distance using a greedy path order optimizer.
\todo{[implementation needed!]}

>>>>

\begin{enumerate}
\item Compute a skeletonization of the outline shape based on the Voronoi diagram.
\item Use a significance measure to mark portions of the skeleton which lie in the center.
\item Determine the bead count values in the marked regions and handle transitions to different bead count values.
\item Generate junctions along unmarked bones of the skeleton and connect the junctions together to form the toolpaths.
\end{enumerate}

@@

\subsubsection{Significance measure}\label{sec:significance_measure}
Places where the naive concentric toolpathing strategy would create large underfilling and overfilling areas are locations which intuitively look like `the center' of the shape.
This can be formalized by looking at the significance measure knows as the \emph{bisector angle}.
The bisector angle $\alpha$ is the angle $\angle{p_ovp_1}$ between a location $v$ on a bone of the skeleton and its two supporting polygon points $p_0$ and $p_1$. \cite{attali1996modeling}
For a polygon with a pointy wedge area of an angle $\beta$, we have $\alpha = 180\degree - \beta$, which corresponds to overfill areas and underfill areas the size of $\nicefrac12 w^2 \left( \nicefrac14 \tan ( \alpha / 2) - \alpha / 2 \right)$ when filled using a naive strategy of constant bead width $w$.
See \cref{naive_overfill_underfill}.
The bisector angle is therefore an exact measure of the amount of overfill and underfill.
Contrary to related literature we will \emph{keep} the non-significant regions of the skeleton.
Our framework will employ a different toolpathing strategy to prevent overfill and underfill areas if they would be so large that $\alpha > \alpha_\text{max}$.

@@

A computationally efficient way to compute $\alpha$ can be obtained by looking at the ratio between feature radius $R$ and the Euclidean distance.
if $ | R(v_1) - R(v_0) | / |v_1 - v_0| >  \cos(\alpha_\text{max} / 2)$ then $\alpha > \alpha_\text{max}$.
See \cref{distance_based_angles}.
This ratio can readily be applied to simple skeleton segments to determine whether they are significant.
For skeletal segments generated by a polygon vertex at $(0,d)$ and a line segment through $(0,0)$ and $(0,\epsilon)$ or another vertex $(0,0)$, we can determine the significant portion by evaluating $\frac{\partial R}{\partial x} > \cos(\alpha_\text{max} / 2)$.
For parabolas we introduce extra points in the discretization at the two bounds where $| x | = d  / \tan(\alpha_\text{max} / 2)$
and for bones with vertex-vertex support we introduce discretization points at the two bounds where $| x | = \nicefrac12 d  / \tan(\alpha_\text{max} / 2)$.
This ensures that the discretization doesn't alter the extent and existence of significant regions.

@@

\subsubsection{Bead counts}
In order to apply the beading strategy to the VQ
we assign each marked node $v$ of the skeleton a bead count $v.B^*$ which is initialized to $b(R(v))$, but it is modified in order to prevent discontinuities and remaining gaps.

We also assign a bead count $v.B^*=b(R(v))$ to all nodes $v$ which are local maxima, i.e. the $v$ for which $\forall v_\text{n} \in v.\text{neighborhood} : R(v) > R(v_\text{n})$. 
This prevents the single overfill or underfill region in unmarked regions.
See \cref{rounded_vs_unrounded}.

\subsubsection{Transitioning}
In order to prevent discontinuities in the toolpaths around locations where the bead count changes, we need a way to smoothly transition to a different number of beads around locations where the 3D model has a changing radius.
See \cref{transitions}.
The transition will be placed anchored on each location $v$ where $R(v) = \nicefrac12 b^{-1}(n+1)$ for any lower bead count $n$.
At the start $v_0$ and end $v_1$ of the transition we introduce a new node in the skeleton and connect it to the outline as per the VQ requirement.
We then assign $v_0.B^* = n$ and $v_1.B^* = n + 1$.
If there is any node $v_x$ in between $v_0$ and $v_1$ we assign it a fractional bead count by linear interpolation: $v_x.B^* = n + D(v_0, v_x)/t(n)$, where $D(.)$ is the Euclidean distance measured along the marked bone(s).
See \cref{distance_rounding_transition}.

@@

\subsubsection{Beading conflicts}
In marked regions of the VQ we employ the approach of applying the optimal bead counts and employing transitioning, while in unmarked regions the beading is only governed by the beading of locations with locally optimal feature radius.
Marked bones always form chains and don't branch out because of the angle constraint on $\alpha_\text{max}$. \todo{Not true anymore}
At the end of such chain $v$ we may need to transition from the former approach to the latter.
This only happens when the feature radius continues increasing: if $\exists v_n \in v.\text{neighborhood}: R(v) < R(v_n) \land | R(v) - R(v_n) | / |v - v_n| >  \cos(\alpha_\text{max} / 2)$ then the bead count assigned by the marked region doesn't fit with a beading which would follow from a local optimum at $v_n$ or some farther node.
See \cref{beading_conflict_problem}.

@@

\begin{figure}
\centering
\includegraphics[width=.5\columnwidth]{sources/method/transition_to_insignificance_problem.pdf}
\caption{
Problems at the boundary of marked regions highlighted with red ovals.
The beading approach from the marked region (blue) results in toolpaths (black) which don't line up with the approach following from the beading which results from the bead count at the local maximum in the bottom center.
}
\label{beading_conflict_problem}
\end{figure}

The transition length of these transitions to unmarked regions is determined by the fractional difference in bead count between the two beadings:
if $v.B^* < n$ then $t = t(n) . (R(v) - v.L_n)$
otherwise $t = t(n) . (1 - (R(v) - v.L_n))$,
where $n = \argmax_i v.L_i < R(v)$.
At a distance $t$ from $v$ along marked bones we introduce an new node $v_0$ for the end of the transition and again we connect it to the outline as per the VQ requirements.

We modify $v.B^*$ such that it coincides with the beading propagated to $v$ from some farther local optimum or another marked region.
(This propagation is explained in \cref{section_toolpath_extraction}.)
We assign $v.B^* = n + 1 - \epsilon$, so that the transition switches from the marked beading to the beading of the unmarked approach entirely.
For nodes $v_x$ in between $v_0$ and $v$ we interpolate $v_x.B^*$ again linearly based on the Euclidean distance along the bones in between $v$ and $v_0$.

@@

The angle requirement ensures that there is never a 3-way junction of significant bones. \todo{not true anymore}

@@

Now that we have the skeleton set up, we determine the locations along the bones where the vertices of the toolpath polygons should be: the \emph{junctions}.
For marked regions we have already inroduced joints around the transitions which serve as those locations.
For unmarked regions we can simply apply the beading strategy in order to find the toolpath verts.

This needs to be stable against small perturbations in the outline polygon.
See \cref{heterogeneous_joint_generation}.
For any sequence on unmarked bones from the outline polygon inward we define the stable end point as the first vertex which is connected to a marked VQ segment.
Based on the feature radius $R$ there we place junctions along the unmarked bones following the beading strategy.

@@

\subsubsection{Toolpath extraction algorithm}
After the beading strategy has been applied we have a beading associated with each marked node as well as each node with locally optimal $R$.
In order to determine the junction locations of quads which don't contain any such node, we propagate the beading information down.
We therefore sort all quads on their highest $R$-value and process the high $R$ quads first.
The beading information is then passed down along the bones of the quad to lower $R$ nodes if those nodes don't already have a beading associated.

@@

When all extrusion segments are generated, we combine them into semi-continuous polyline toolpaths.
These polylines generally form closed polygons, but at marked regions of the VQ the toolpaths consist of single bead polyline toolpaths.
See \cref{single_bead_strategy}.
These single bead polylines can connect to polygonal toolpaths and they can also meet each other at 3-way junctions.
We determine the printing order for all segments of a given bead index by minimizing the travel distance using a greedy path order optimizer.
\todo{[implementation needed!]}
